---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, message=FALSE, warning=FALSE}
###import data
activity <- read.csv(file = "activity.csv", stringsAsFactors = F)

### format date
library(lubridate)
activity$date <- ymd(activity$date)
```



## What is mean total number of steps taken per day?

Histogram of steps looks like these:

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(activity) + geom_histogram(aes(x=steps), binwidth = 5)

```
Mean of steps taken each day is `r mean(activity$steps, na.rm=T)` and median is `r median(activity$steps, na.rm=T)`

## What is the average daily activity pattern?


```{r, message=FALSE, warning=FALSE, include=FALSE}
AvgSteps <-  activity%>%group_by(interval)%>%summarise(Avg.Steps=mean(steps,na.rm = T))%>%arrange(desc(Avg.Steps))

```
Daily 5-minute interval with maximum steps:
```{r}
AvgSteps[1,]
```

Distribution of average number of steps per 5 minute interval.

```{r}
library(dplyr)
activity%>%group_by(interval)%>%summarise(Avg.Steps=mean(steps,na.rm = T))%>%
        ggplot() + geom_line(aes(x=interval,y=Avg.Steps))
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
